SUPPLEMENTARY INFORMATION: NMRlipids III: Lipid-cholesterol interactions in atomistic resolution molecular dynamics simulations

### 2019-05-07 MANUSCRIPT/manuscript.tex
Update manuscript

Addition of cholesterol substantially increases the acyl chain C-H order parameters in PC lipid bilayers (Fig. \ref{OrderParametersCHOL}),
which can be explained by increased trans conformations in acyl chains \cite{ferreira13,??}.
%Acyl chain order parameters for pure 1-palmitoyl-2-oleoylphosphatidylcholine (POPC) bilayer
%and mixture with 50 mol\% of cholesterol from different simulations and experiments
%are shown in Fig. \ref{OrderParametersCHOL}.
The absolute values of order parameters of sn-1 chain from $^{13}$C NMR experiments \cite{ferreira13}
exhibit approximately linear increase with cholesterol up to equimolar mixture (Fig. \ref{slopes}), which is expected because
phase separation is not observed for this mixture \cite{ionova12,ferreira13}.
%thus the intermolecular interactions are expected to have maximum effect in equimolar mixture.

Experimental acyl chain order parameters are typically well reproduced in the state of the art lipid models
for pure lipid bilayers (for review see \cite{ollila16}). This is also mainly observed in the data
presented here, except that CHARMM36 slightly overestimates the order parameters and MACROG model
underestimates order parameters in the beginning of sn-1 chain. In equimolar mixtures with cholesterol
all models overestimate the order parameters. For Berger/H\"oltje and MACROG models the overestimation is significant,
while less severe for CHARMM36 and Slipid models.\\
\todo{Is overestimation in CHARMM36 and Slipid significant or not?} \\
\todo{Lipid14 results from recent article should be added in this figure} 
\begin{figure*}[]
  \centering
  \includegraphics[width=17.2cm]{../FIGS/OrderParametersCHOL.eps}
  \caption{\label{OrderParametersCHOL}
    Order parameters from simulations and experiments for acyl chains of  1-palmitoyl-2-oleoylphosphatidylcholine (POPC).
  }
\end{figure*}

\begin{figure*}[]
  \centering
  \includegraphics[width=17.2cm]{../FIGS/OrderParametersCHOLfitness.eps}
  \caption{\label{OrderParametersCHOL}
    Difference between order parameters from simulations and experiments for acyl chains of  1-palmitoyl-2-oleoylphosphatidylcholine (POPC).
  }
\end{figure*}

The order parameter changes as a function of cholesterol for each segment are shown in Fig.~\ref{OrderParametersCHOLchanges}
(currently only sn-1). \\
\todo{This figure should be finished and similar for sn-2 should be done as well}
\begin{figure}[]
  \centering
  \includegraphics[width=8cm]{../FIGS/slopes.eps}
  \caption{\label{OrderParametersCHOLchanges}
    Slopes of order parameters as a function of cholesterol.
    %Order parameter changes from simulations and experiments for each segment in sn-1 chain of  1-palmitoyl-2-oleoylphosphatidylcholine (POPC) as a function of cholesterol concentration.
  }

\end{figure}

@@

>>>>

Addition of cholesterol substantially increases the acyl chain C-H order parameters in PC lipid bilayers (Fig. \ref{OrderParametersCHOL}),
which can be explained by increased trans conformations in acyl chains \cite{ferreira13,??}.
%Acyl chain order parameters for pure 1-palmitoyl-2-oleoylphosphatidylcholine (POPC) bilayer
%and mixture with 50 mol\% of cholesterol from different simulations and experiments
%are shown in Fig. \ref{OrderParametersCHOL}.
The absolute values of order parameters of sn-1 chain from $^{13}$C NMR experiments \cite{ferreira13}
exhibit approximately linear increase with cholesterol up to equimolar mixture (Fig. \ref{slopes}), which is expected because
phase separation is not observed for this mixture \cite{ionova12,ferreira13}.
%thus the intermolecular interactions are expected to have maximum effect in equimolar mixture.
\todo{Is overestimation in CHARMM36 and Slipid significant or not?} \\
\todo{Lipid14 results from recent article should be added in this figure} 
\begin{figure*}[]
  \centering
  \includegraphics[width=17.2cm]{../FIGS/OrderParametersCHOL.eps}
  \caption{\label{OrderParametersCHOL}
    Order parameters from simulations and experiments for acyl chains of  1-palmitoyl-2-oleoylphosphatidylcholine (POPC).
  }
\end{figure*}

Experimental acyl chain order parameters of pure POPC lipid bilayers
are well reproduced by most force fields (Fig. \ref{OrderParametersCHOL}),
except for the C$_2$ carbon in {\it sn}-1 chain,
which is typically the case in the state of the art lipid force fields (for review see \cite{ollila16}).
%This is also mainly observed in the data
%presented here, except that
However, the order parameters in the beginning of {\it sn}-1 chain are underestimated in the
MACROG simulation and CHARMM36 simulations ran also with other than Gromacs
(section \ref{CHARMMprograms} in the supplementary information) slightly overestimates the order parameters.
The absolute values of {\it sn}-1 acyl chain order parameters exhibit approximately linear increase upon addition
of cholesterol also in simulations below equimolar mixture (Figs. \ref{slopesberger}-\ref{slopesslipids310K}).
However, the slope of the increase (Fig. \ref{OrderParametersCHOLchanges}) and order parameters
in equimolar mixtures (Fig. \ref{OrderParametersCHOL}) are overestimated by all force fields.
%For Berger/H\"oltje and MACROG models the overestimation is significant,
%while less severe for CHARMM36 and Slipid models.
The overestimation of order parameters observed in CHARMM without cholesterol and in all force fields
with cholesterol is smaller than the contribution by undulations in large simulations (section \ref{undulations} in the supplementary information).
\begin{figure}[]
  \centering
  \includegraphics[width=8cm]{../FIGS/slopes.eps}
  \caption{\label{OrderParametersCHOLchanges}
    Slopes of order parameters as a function of cholesterol.
  }
\end{figure}

@@

\section{Presenting the quality of lipid bilayer MD simulation with binary mixtures}

\begin{figure*}[]
  \centering
  \includegraphics[width=17.2cm]{../FIGS/OrderParametersCHOLfitness.eps}
  \caption{\label{OrderParametersCHOLfitness}
    Difference between order parameters from simulations and experiments for acyl chains of  1-palmitoyl-2-oleoylphosphatidylcholine (POPC).
  }
\end{figure*}

### 2019-05-07 MANUSCRIPT/manuscriptSUPPL.tex
Update manuscript

\section{CHARMM36 results from different simulation packages}
The results from CHARMM36 model for lipid bilayers from different 
simulation packages have been reported to give different results in
the literature \cite{piggot12,lee16}. The results are mainly
dependent on different Lennard-Jones cut-off settings, but
all the details are not quite understood. In this work we use
the results from Gromacs 5 with settings suggested to be optimal
by Gromacs webpage.~\todo{List these settings here or in SI.} We also compared the results from Gromacs 5 with
these settings to the results simulated with NAMD, OpenMM and literature
values. Based on comparison shown in Fig. \ref{programsCOMP}, we conclude
that Gromacs 5 with settings suggested in webpage gives consistent
results with the literature and other simulation packages. Thus these
results are used in the main body of the paper. However, order parameters
are slightly overestimated respected to the experiments also with these
settings. \\
\todo{Should be extend this discussion based on discussion at  https://github.com/NMRLipids/NmrLipidsCholXray/issues/4 ?}
 \begin{figure}[]
  \centering
  \includegraphics[width=8cm]{../FIGS/OrderParametersPROGRAMS.eps}

@@

>>>>

\section{CHARMM36 results from different simulation packages} \label{CHARMMprograms}
The results from CHARMM36 model for lipid bilayers from different 
simulation packages have been reported to give different results in
the literature \cite{piggot12,lee16}. The results are mainly
dependent on different Lennard-Jones cut-off settings, but
all the details are not quite understood. In this work we use
the results from Gromacs 5 with settings suggested to be optimal
by Gromacs webpage.~\todo{List these settings here or in SI.} We also compared the results from Gromacs 5 with
these settings to the results simulated with NAMD, OpenMM and literature
values. Based on comparison shown in Fig. \ref{programsCOMP}, we conclude
that Gromacs 5 with settings suggested in webpage gives consistent
results with the literature and other simulation packages. Thus these
results are used in the main body of the paper. However, order parameters
are slightly overestimated respected to the experiments also with these
settings. \\
\todo{Should be extend this discussion based on discussion at  https://github.com/NMRLipids/NmrLipidsCholXray/issues/4 ?}
 \begin{figure}[]
  \centering
  \includegraphics[width=8cm]{../FIGS/OrderParametersPROGRAMS.eps}

@@

 \section{Effect of undulations on order parameters}
 \todo{This should be written based on contributions in this issue:
   https://github.com/NMRLipids/NmrLipidsCholXray/issues/16
 }
